---
title: "Assignment Wind Power"
output:
  pdf_document: default
  html_notebook: default
---

# Descriptive statistics 

### 1. Read the data tuno.txt into R.
```{r}
data <- read.delim("tuno.txt",header = TRUE, sep = " ", dec = ".")
head(data,6)
```


### 2. Make a graphical presentation of data or parts of the data, and present some summary statistics.

#### Summary statistics:
```{r}
summary(data[c(0,4:6)],digits = 2)
```

#### Graphical presentation:
```{r}
# Distribution of the power production

library(ggpubr)

gghistogram(data, x = "pow.obs", bins = 30, 
             add = "mean")

ggboxplot(data, y = "pow.obs", width = 0.5)

plot(data$r.day,data$pow.obs)
```



```{r}
# do we in some months produce more power?
library(tidyverse)

mean_by_month_power <- data %>% 
  group_by(month) %>% 
  summarize(averaged.power = mean(pow.obs))

plot(mean_by_month_power)

# Are some months on average more windy?
mean_by_month_wind <- data %>% 
  group_by(month) %>% 
  summarize(averaged.wind= mean(ws30))

plot(mean_by_month_wind)

```

```{r}
#does a high wind prediction equal a high power production?
# Indeed yes!
plot(data$ws30,data$pow.obs)
```
```{r}
# correlation matrix with power, wind speed and wind direction:
cor(data)
```

#### Normalize the power production before modelling:
```{r}
normalize <- function(x) {
  return ((x+0.01 - min(x)) / (max(x) - min(x)+0.04))}

data$pow.obs.norm<-normalize(data$pow.obs)

head(data,6)

```

### 3. Simple models:
```{r}
data$pow.obs.norm.log<-log(data$pow.obs.norm)

qqnorm(data$pow.obs.norm, pch = 1, frame = FALSE)
qqline(data$pow.obs.norm, col = "steelblue", lwd = 2)
```
The curvature of the qq-plot indicates that a beta distribution might fit the data well.
**http://onlinestatbook.com/2/advanced_graphs/q-q_plots.html


#### Fitting a beta distribution to the data.
The likelihood function for the wind data, can be defined as a sum of the log density functions $$l(\theta)=\sum_{i=1}^{n}log(p_{\theta}(x_{i}))$$

where $p_{\theta}$ in this case is the beta density function
given as:
$$f(x) = \frac{(x-a)^{p-1}(b-x)^{q-1}}{B(p,q) (b-a)^{p+q-1}}
\hspace{.3in} a \le x \le b; p, q > 0$$

where p and q are the shape parameters, a and b are the lower and upper bounds, respectively, of the distribution, and B(p,q) is the beta function. The beta function has the formula
$$B(\alpha,\beta) = \int_{0}^{1} {t^{\alpha-1}(1-t)^{\beta-1}dt}$$
The likelihood function can easily be implemented in R:
```{r}
Beta.likelihood=function(shape1, shape2){ -sum(dbeta(x=data$pow.obs.norm,shape1=shape1, shape2=shape2, log=TRUE))}
```

Now the optimal value of the two shape parameters must be found. This can be done using profile likelihood, the profile likelihood is defined as:
Let $(\theta, \delta)$ be the full parameter and $\theta$ the parameter of interest. Given the joint likelihood $L(\theta, \delta)$ the profile likelihood of $\theta$ is 
$$L(\theta)=max_{\delta}L(\theta, \delta)$$
Where the maximation is performed at fixed value of $\theta$ .

```{r}
#Profile likelihood for shape1 
shape1=seq(0.35, 0.68,by=0.001)
shape2=seq(0.8, 1.6,by=0.01)


lp.shape1= function(shape1){
  ## Function for inner optimization:
  fun.tmp<- function(shape1, shape2){
    Beta.likelihood(shape1,shape2)
  }
  # Find the profile likelihood
  optimise(fun.tmp,c(0.1,10), shape1=shape1)$objective
}

(opt.shape1=optimise(lp.shape1, c(0.1,10))$minimum)


```

```{r}
#Profile likelihood for shape2

lp.shape2= function(shape2){
  ## Function for inner optimization:
  fun.tmp<- function(shape1, shape2){
    Beta.likelihood(shape1,shape2)
  }
  # Find the profile likelihood
  optimise(fun.tmp,c(0.1,10), shape2=shape2)$objective
}

(opt.shape2=optimise(lp.shape2, c(0.1,10))$minimum)
```

This means, that the MLE for the two shape parameters are $shape1=0.5051217$ and $shape2=1.147125$.

```{r}

# Profile likelihoods:
lp.shape1.plot= sapply(shape1, lp.shape1)
lp.shape2.plot= sapply(shape2, lp.shape2)

par(mfrow=c(2,1), mar=c(3.3,3.3,1,1),mgp=c(2,0.7,0))
plot(shape1,exp(-lp.shape1.plot), type="l", xlab="Profile Likelihood Shape 1")
lines(c(-2,10),exp(-qchisq(0.95,df=1)/2)*c(1,1),col=2)

plot(shape2,exp(-lp.shape2.plot), type="l", xlab="Profile Likelihood Shape 2")
lines(c(-2,10),exp(-qchisq(0.95,df=1)/2)*c(1,1),col=2)

```
Plot distribution of power.obs and the optimal fit for beta distribution.
```{r}
# Add a Beta Curve
x <- data$pow.obs.norm
h<-hist(x, xlab="Power Obs Norm",
   main="Histogram with Beta Curve")
xfit<-seq(min(x),max(x),length=40)
yfit<-dbeta(xfit,opt.shape1,opt.shape2)
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2)
```
AIC is calculated using the formular:
$$AIC=2k-2ln(\hat{L})$$
where L is the maximum value for the likelihood function for the model. 
```{r}
# Maximum value of the likelihood function for the model
(AIC.beta =2*2-2*log(-Beta.likelihood(opt.shape1,opt.shape2)))
```


#### Fitting a gamma distribution to the data.

The density function for a gamma distribution is: 
$$f(x) = \frac{(\frac{x-\mu}{\beta})^{\gamma - 1}\exp{(-\frac{x-\mu}
{\beta}})} {\beta\Gamma(\gamma)}  \hspace{.2in}  x \ge \mu; \gamma,
\beta > 0$$

```{r}
Gamma.likelihood=function(shape, rate){ -sum(dgamma(x=data$pow.obs.norm,shape=shape, rate=rate, log=TRUE))}

```

Now the optimal value of the two shape parameters must be found. This can be done using profile likelihood. 

```{r}
#Profile likelihood for shape 
shape=seq(0.1, 0.9,by=0.001)
rate=seq(0.1, 0.9,by=0.01)


lp.gamma.shape= function(shape){
  ## Function for inner optimization:
  fun.tmp<- function(shape, rate){
    Gamma.likelihood(shape,rate)
  }
  # Find the profile likelihood
  optimise(fun.tmp,c(0.1,10), shape=shape)$objective
}

(opt.gamma.shape=optimise(lp.gamma.shape, c(0.1,10))$minimum)
```
Find the optimal value for the rate parameter. 

```{r}
#Profile likelihood for rate 

lp.gamma.rate= function(rate){
  ## Function for inner optimization:
  fun.tmp<- function(shape, rate){
    Gamma.likelihood(shape,rate)
  }
  # Find the profile likelihood
  optimise(fun.tmp,c(0.1,10), rate=rate)$objective
}

(opt.gamma.rate=optimise(lp.gamma.rate, c(0.1,10))$minimum)
```

Plot distribution of power.obs and the optimal fit for beta distribution.
```{r}
# Add a Gamma Curve
x <- data$pow.obs.norm
h<-hist(x, xlab="Power Obs Norm",
   main="Histogram with Gamma Curve")
xfit<-seq(min(x),max(x),length=40)
yfit<-dgamma(xfit,opt.gamma.shape,opt.gamma.rate)
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2)
```
Calculating AIC for the gamma fit. 

```{r}
# Maximum value of the likelihood function for the model
k = 2 #Number of parameters
(AIC.gamma =2*k-2*log(-Gamma.likelihood(opt.gamma.shape,opt.gamma.rate)))
```
#### Fitting a log-normal distribution to the data.

The density function for a gamma distribution is: 
$$\frac{1}{x \sigma \sqrt{2 \pi}} \exp \left(-\frac{(\ln x-\mu)^{2}}{2 \sigma^{2}}\right)$$

```{r}
LogNormal.likelihood=function(meanlog, sdlog){ -sum(dlnorm(x=data$pow.obs.norm,meanlog=meanlog, sdlog=sdlog, log=TRUE))}
```

Profile likelihood functions for the log-normal distribution. 

```{r}
#Profile likelihood for meanlog
meanlog=seq(0.1, 0.9,by=0.001)
sdlog=seq(0.1, 0.9,by=0.01)


lp.LogNormal.meanlog= function(meanlog){
  ## Function for inner optimization:
  fun.tmp<- function(meanlog, sdlog){
    LogNormal.likelihood(meanlog,sdlog)
  }
  # Find the profile likelihood
  optimise(fun.tmp,c(0.1,10), meanlog=meanlog)$objective
}

(opt.LogNormal.meanlog=optimise(lp.LogNormal.meanlog, c(0.1,10))$minimum)
```
Find the optimal value for sdlog. 

```{r}
#Profile likelihood for sdlog
lp.LogNormal.sdlog= function(sdlog){
  ## Function for inner optimization:
  fun.tmp<- function(meanlog, sdlog){
    LogNormal.likelihood(meanlog,sdlog)
  }
  # Find the profile likelihood
  optimise(fun.tmp,c(0.1,10), sdlog=sdlog)$objective
}

(opt.LogNormal.sdlog=optimise(lp.LogNormal.sdlog, c(0.1,10))$minimum)
```
Plot distribution of power.obs and the optimal fit for log-normal distribution.
```{r}
# Add a Log-normal Curve
x <- data$pow.obs.norm
h<-hist(x, xlab="Power Obs Norm",
   main="Histogram with Log-normal Curve")
xfit<-seq(min(x),max(x),length=40)
yfit<-dlnorm(xfit,opt.LogNormal.meanlog,opt.LogNormal.sdlog)
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2)
```

Calculate AIC for log-normal distribution. 
```{r}
# Maximum value of the likelihood function for the model
k=2
(AIC.LogNormal =2*k-2*log(-LogNormal.likelihood(opt.LogNormal.meanlog,opt.LogNormal.sdlog)))
```








