---
title: Assignment - Financial Data
output: html_notebook
---

## 1: Descriptive statistics and simple models
```{r}
setwd("/Users/akterminsprove/Desktop/DTU/5. Semester/02418 Statistisk Modellering Teori og Anvendelser/Assignments/Financial Data")
data <- read.csv("finance_data.csv", header=TRUE,sep=";")
data$time=as.Date(data$time)
head(data,6)
```

### a) Present the data, estimate the parameters in a normal model, and asses if the normal model is appropriate.

A scatter plot and a qq-plot of the data is made:
```{r}
plot(data$time,data$SLV)
qqnorm(data$SLV, pch = 1, frame = FALSE)
qqline(data$SLV, col = "steelblue", lwd = 2)
```
It seems reasonable that the data could fit a normal distribution, though some deviations at the ends.

Ignoring irrelevant constant terms the log-likelihood of a normal distribution can be formulated as:
$$ l(\mu,\sigma^2)=\frac{-n}{2} log(\sigma^2)-\frac{1}{2\sigma^2}\cdot \sum_{i}(x_i-\mu)^2$$
Where the maximum likelihood estimates are:
$$\hat{\mu}= \bar{x}$$
$$\hat{\sigma^2}=\frac{1}{n}\cdot \sum_i (x_i-\bar{x})^2$$
```{r}
normal.likelihood=function(mu,sigma2){
  -sum(dnorm(x=data$SLV,mu,sigma2,log=TRUE))}

```


```{r}
#Profile likelihood for mu 
mu=seq(1, 10,by=0.0001)
sigma2=seq(0.01, 5,by=0.01)


lp.mu= function(mu){
  ## Function for inner optimization:
  fun.tmp<- function(mu, sigma2){
    normal.likelihood(mu,sigma2)
  }
  # Find the profile likelihood
  optimise(fun.tmp,c(0.001,10), mu=mu)$objective
}

(opt.mu=optimise(lp.mu, c(0.001,10))$minimum)

```

```{r}
#Profile likelihood for shape1 
mu=seq(1, 10,by=0.0001)
sigma2=seq(0.01, 5,by=0.01)


lp.sigma2= function(sigma2){
  ## Function for inner optimization:
  fun.tmp<- function(mu, sigma2){
    normal.likelihood(mu,sigma2)
  }
  # Find the profile likelihood
  optimise(fun.tmp,c(0.001,10), sigma2=sigma2)$objective
}

(opt.sigma2=optimise(lp.sigma2, c(0.001,10))$minimum)
```

```{r}
normal.likelihood(opt.mu,opt.sigma2)
```

```{r}
x=data$SLV
h<-hist(x, xlab="Power Obs Norm",
   main="Histogram with Normal Curve")
xfit<-seq(min(x),max(x),length=40)
yfit<-dnorm(xfit,opt.mu,opt.sigma2)
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2)
```
AIC function:
```{r}
AIC=function(nll,k){
  2*k-2*(-nll)
}


(AIC.normal=AIC(normal.likelihood(opt.mu,opt.sigma2),2))
```


#### b) Hypothesize a model that could fit the data better (Hint: consider tail probabilities), and compare with the normal model estimated above

A model that considers the tail probabilities is a t-distribution with $v$ degrees of freedom, if $v=1$ we obtain the cauchy model, while $v \rightarrow \infinity$ gives the normal distribution
```{r}
mean
```

```{r}
library(metRology)
t.likelihood=function(params){
  -sum(dt.scaled(x=data$SLV, df=params[1], mean = params[2], sd = params[3], log = TRUE))}
```

```{r}
(opt.t=nlminb(c(1,1,1),t.likelihood))
```

```{r}
x=data$SLV
h<-hist(x, xlab="Power Obs Norm",
   main="Histogram with t-curve Curve")
xfit<-seq(min(x),max(x),length=40)
yfit<-dt.scaled(xfit, df=opt.t$par[1],mean = opt.t$par[2],sd =opt.t$par[3])
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2)
```

```{r}
(AIC.t=AIC(t.likelihood(opt.t$par),3))
```

#### c) represent the final model (i.e. relevant keynumbers for the estimates)

 confidence intervaller på parameterne etc.. 
