---
title: "Assignment Survival Data"
output:
  pdf_document: default
  html_notebook: default
---

#Analysis of binary data

###  Read the Logistic.txt into R.

```{r}
data <- read.delim("Logistic.txt",header = TRUE, sep = "", dec = ".")
data$AIDS_no=data$n-data$AIDS_yes
head(data,2) 

n <- sum(data[,3]) #Number of patient #n 
k <- sum(data[,2]) #Number of patients with AIDS #Sucesses #k
theta <- s/n #Probability that you have Aids
```
The data consist of patients being treated with AZT and the number of patients with AIDS, as well as patients not being treated with AZT and the number of patients having AIDS. It looks like more patients get AIDS if they are not treated with AZT.

### Fitting data to binomial distribution (i.e. consider all data as coming from the same population)
The log-likelihood of the binomial distribution is:
$$l(\theta)=\sum_{k=0}^{n}k\cdot log(\theta)+(n-k)\cdot log(1-\theta)+c(n,k)$$
where n is the number of patients, k is the number of patient with AIDS. Since it is considered as coming from the same population i.e. follow the same distibution, the data is fitted to one binomial distribution.

```{r}

#Formulating the likelihood function
binomial.log.likelihood = function(theta,success,size){
  -(sum(dbinom(x=success,size=size,prob=theta, log=TRUE)))}

(theta.opt=optimize(binomial.log.likelihood, c(0.01,0.99),success=sum(data$AIDS_yes),size=sum(data$n))$minimum)

(nlminb(start=0.5,objective=binomial.log.likelihood,lower=0,success=k,size=n)$par)

```

### Fit the Binomial separately to the two distributions and test if there is a difference between the groups.

```{r}
#Theta for patients treated with AZT. This is the probability of having AIDS when treated with AZT. 
#(theta_AZT_yes.opt=optimize(binomial.log.likelihood, c(0.01,0.99),x=data$AIDS_yes[1],size=data$n[1])$minimum)

(theta_AZT_yes.opt=nlminb(start=0.5,binomial.log.likelihood,lower=0,success=data$AIDS_yes[1],size=data$n[1])$par)

(theta_AZT_no.opt=nlminb(start=0.5,binomial.log.likelihood,lower=0,success=data$AIDS_yes[2],size=data$n[2])$par)

#Theta for patients treated without AZT. This is the probability of having AIDS without treament. 
#(theta_AZT_no.opt=optimize(binomial.log.likelihood, c(0.01,0.99),x=data$AIDS_yes[2],size=data$n[2])$minimum)
```
As expected, the probability for getting AIDS is lower when treated with AZT. 


**Finding the confidence interval for the probabilities.** 


```{r}
k1 <- data$AIDS_yes[1] #ATZ_yes
k2 <- data$AIDS_yes[2] #ATZ_no
n1 <- data$n[1]
n2 <- data$n[2]
p1 <- k1/n1
p2 <- k2/n2
```


```{r}
library(numDeriv)
I <- hessian(binomial.log.likelihood,theta,success=k,size=n)
I_AZT_yes <- hessian(binomial.log.likelihood,theta_AZT_yes.opt,success=k1,size=n1)
I_AZT_no <- hessian(binomial.log.likelihood,theta_AZT_no.opt,success=k2,size=n2)

#The standard error 
se <- 1/sqrt(I)
se_AZT_yes <- 1/sqrt(I_AZT_yes)
se_AZT_no <- 1/sqrt(I_AZT_no)



#confidence intervallet
(p+c(-1,1)*qnorm(0.975)*se/sqrt(n))
(p1+c(-1,1)*qnorm(0.975)*se_AZT_yes/sqrt(n1))
(p2+c(-1,1)*qnorm(0.975)*se_AZT_no/sqrt(n2))

```
Now the parameters and their corresponding CI are found. 

The null hypothesis is that the two distributions are equal meaning that if AZT_yes follows $B(m,\pi_x)$ and AZT_no follows $B(n,\pi_y)$ $\pi_x=\pi_y$.
```{r}
x=data$AIDS_yes[1]
y=data$AIDS_no[1]
m=sum(data$AIDS_yes)
n=sum(data$AIDS_no)
N=m+n
t=x+y

#The standard test of equality of proportions is the famous chi-squared test
(chi.squared.test=N*(x*(n-y)-y*(m-x))^2/(m*n*t*(N-t)))

(pchisq(chi.squared.test, df=1, lower.tail=FALSE))
```
With a p-value of 0.008816159 we have strong evidence against the null hypothesis meaning that the two groups does not follow the same distribution.

To proceed with a likelihood analysis, the number of people with AIDS (X) in the group treated with AZT is $B(m,\pi_x)$ and independently the number of people with AIDS (Y)  in the second group is $B(n,\pi_y)$.

On observing x and y the joint likelihood of $(\pi_x,\pi_y)$:
$$L(\pi_x, \pi_y)=\pi_x^x*(1-\pi_x)^{m-x}\pi_y^y(1-\pi_y)^{n-y}$$

The log odds-ratio is defined by:
$$\theta=\log \frac{\pi_{x} /\left(1-\pi_{x}\right)}{\pi_{y} /\left(1-\pi_{y}\right)}$$
```{r}
(theta.compare = log((theta_AZT_yes.opt/(1-theta_AZT_yes.opt))/(theta_AZT_no.opt/(1- theta_AZT_no.opt))))

(log((x/(m-x))/(y-(n-y))))

```


The null hypothesis is $H_0: \theta=0$. Since $\theta$ is significantly different from zero, the null-hypothesis is rejected and it is concluded that the groups are significantly different from each other. This means that there is a difference in if a patient receives treament with AZT or not. 


### Estimate parameters in the model (p0 probability of AIDS in control group, p1 probability of AIDS in treatment group) and report a confidence interval for the parameter describing the dif- ference, compare with the result above.

```{r}

k1 <- data$AIDS_yes[1] #ATZ_yes
k2 <- data$AIDS_yes[2] #ATZ_no
n1 <- data$n[1]
n2 <- data$n[2]

#Negative log-likelihood function 

logL <- function(theta){
  
  b0 <- theta[1]
  b1 <- theta[2]
  
 k2*((b0 +b1*0)-log(1+exp(b0+b1*0)))+(n2-k2)*((b0+b1*0)-log(1+exp(b0+b1*0)))+k1*((b0 +b1*1)-log(1+exp(b0+b1*1)))+(n1-k1)*((b0+b1*0)-log(1+exp(b0+b1*1)))
  
}

opt <- nlminb(c(0,0),objective = logL,lower=c(-Inf,-Inf))

```



The formulas look like a binomial model. This binomial model has the log-likelihood
$$ $$
```{r}

#fit.glm =glm(y~x, family=binomial)
```

#Analysis of the survival time data

```{r}

#Load data
data.time <- read.delim("actg320.txt",header = TRUE, sep = "", dec = ".")
head(data.time,2) 
```


```{r}
#Count number of patient without treatment and without AIDS
(sum(data.time$tx == 0 & data.time$event == 0))
#Count number of patient without treatment and AIDS
(sum(data.time$tx == 0 & data.time$event == 1))
#Count number of patient with treatment and without AIDS
(sum(data.time$tx == 1 & data.time$event == 0))
# Count number of patient with treatment and AIDS
(sum(data.time$tx == 1 & data.time$event == 1))

#Calculating proportions

#No aids out of people without treatment
(sum(data.time$tx == 0 & data.time$event == 0))/sum(data.time$tx == 0)
(sum(data.time$tx == 0 & data.time$event == 1))/sum(data.time$tx == 0)
(sum(data.time$tx == 1 & data.time$event == 0))/sum(data.time$tx == 1)
(sum(data.time$tx == 1 & data.time$event == 1))/sum(data.time$tx == 1)
```
### Fit an exponential distribution to all the data. 

```{r}
exp.likelihood=function(rate){
  -sum(dexp(x=data.time$time[data.time$event==1], rate=rate, log=TRUE))- sum(pexp(q=data.time$time[data.time$event==0], lower.tail = FALSE, rate=rate, log=TRUE))
  }


(nlminb(start=0.5,objective = exp.likelihood,lower=0)$par)
(nlminb(start=0.5,objective = exp.likelihood,lower=0)$objective)
#(optimize(exp.likelihood,c(0, 100),maximum=TRUE)$maximum)

```

Fit an exponential distribution to each group. 

```{r}

exp.likelihood_no_treat=function(rate){
  -sum(dexp(x=data.time$time[data.time$event==1 & data.time$tx == 0], rate=rate, log=TRUE))- sum(pexp(q=data.time$time[data.time$event==0 & data.time$tx == 0], lower.tail = FALSE, rate=rate, log=TRUE))
}

(nlminb(start=0.5,objective = exp.likelihood_no_treat,lower=0)$par)
(nlminb(start=0.5,objective = exp.likelihood_no_treat,lower=0)$objective)



exp.likelihood_with_treat=function(rate){
  -sum(dexp(x=data.time$time[data.time$event==1 & data.time$tx == 1], rate=rate, log=TRUE))- sum(pexp(q=data.time$time[data.time$event==0 & data.time$tx == 1], lower.tail = FALSE, rate=rate, log=TRUE))
}

(nlminb(start=0.5,objective = exp.likelihood_with_treat,lower=0)$par)
(nlminb(start=0.5,objective = exp.likelihood_with_treat,lower=0)$objective )
```
It can be seen that the model fits the data better when the data is separated into the two treatment groups. 

### Formulate af model where one parameter indicate the treatment effect

```{r}
# Implement negative log likelihood function

nll.exp <- function(rate){
  
  val1 <- -sum(dexp(x=data.time$time[data.time$event==1],rate=rate[1]+rate[2]*data.time$tx[data.time$event==1],log=TRUE))
  
  val2 <- -sum(pexp(q=data.time$time[data.time$event==0],rate=rate[1]+rate[2]*data.time$tx[data.time$event==0],lower.tail=F,log=TRUE))
  
  val1+val2
}

opt.exp <- nlminb(start=c(0.001,0.005),objective=nll.exp,lower=c(-Inf,-Inf))
(opt.exp$par[2])


#Making a confidence interval

I <- hessian(nll.exp,opt.exp$par)
se <- 1/sqrt(I)
(opt.exp$par+c(-1,1)*qnorm(0.975)*se/sqrt(n))
```
It can be seen that the confidence interval does cover zero which means that the two treatment groups are not significantly different from each other. So even though it was found in the previous question that the model fits that data better when the data is separated, there is no significant difference. 

```{r}
nll.b1 <- function(rate,b1){
  
  val1 <- -sum(dexp(x=data.time$time[data.time$event==1],rate=rate[1]+b1*data.time$tx[data.time$event==1],log=TRUE))
  
  val2 <- -sum(pexp(q=data.time$time[data.time$event==0],rate=rate[1]+b1*data.time$tx[data.time$event==0],lower.tail=F,log=TRUE))
  
  val1+val2
}


profile.b1 <- function(b1){
  opt <- nlminb(c(0),nll.b1,lower=(-Inf),b1=b1)
  b0 <-opt$par[1]
  
  -nll.b1(c(b0),b1)
}
```


```{r}
b1 <- seq(-5,5,0.1)
p = c()

# for (i in 1:length(b1)) {
#   p[i] = exp(profile.b1(b1[i]))
# }
  
  #plot(b1,p/max(p),'l')
  plot(b1,profile.b1(b1))

```

 