#################
## Binary data ##
#################

Read the Logistic.txt into R
```{r}
data <- read.delim("Logistic.txt",header = TRUE, sep = "", dec = ".")
data$AIDS_no=data$n-data$AIDS_yes
head(data,2) 
```

Fit a logistic regression model

```{r}
df <- data.frame("AZT" = c(rep(1, 170), rep(0,168)), "AIDS" = c(rep(1, 25), rep(0, 145), rep(1, 44), rep(0, 124)))

head(df, )
         
length(AZT)
length(AIDS)

mod2 <- survreg(Surv(AIDS) ~ AZT, data = df, dist = "logistic")
summary(mod2)
         
```
Calculating the odds ratio

```{r}
exp(coef(mod2)[2]/mod2$scale)
```
Interpretation: The odds of survival until time t is 0.68 times greater for AZT = 1 than for AZT = 0

Confidence interval

```{r}
confint(mod2)

```
Test the hypothesis of no effect of AZT on AIDS

Likelihood ratio test

Directly from R output??

```{r}

```



The Wald test

The null hypothesis of log(scale) = 0 this can be seen directly from the R output
The null hypothesis is rejected on all levels as the p-value is 2e-16 



```{r}

```






#######################
## Suvival time data ##
#######################

Reading the data into R

```{r}
data.time <- read.delim("actg320.txt",header = TRUE, sep = "", dec = ".")
head(data.time,2) 
```

Number of patients getting AIDS or dying in the two groups
```{r}
# Treated group
(sum(data.time$event == 1 & data.time$tx == 1 ))


# Non treated group
(sum(data.time$event == 1 & data.time$tx == 0 ))

```
The total follow up time
follow up time = sum of the time that you follow all of the participants
```{r}
print('Follow up for new treatment group')
(sum(data.time$time[data.time$tx == 0])/365)


print('Follow up for new treatment group')
(sum(data.time$time[data.time$tx == 1])/365)
```
Plotting survival function in the two treatment groups

```{r}
library(survival)

Surv.ByTreat <- survfit(Surv(time, event == 1) ~ tx,conf.type = "log-log",data = data.time)

plot(Surv.ByTreat, conf.int = T, las = 1, xlab = "Days",ylab = "Estimated Survival Prob.", col=2:3, lwd = 2, ylim=c(0.8,1))
legend("bottomleft", col=2:3, c("Control","Treated"), lwd=2)


```
From the above plot it seems that the treated group is doing best

Plot the cumulative incidence functions

```{r}
plot(Surv.ByTreat, fun=function(x) { 1- x }, conf.int = F,las = 1, xlab = "Days",ylab = "Estimated Failure Prob.", col=2:3, lwd = 2)
legend("topleft", col=2:3, c("Control","Treated"), lwd=2)

# I would prefer the survival function as it is more positive to think about a survival probability than it is about a failure probability

```

Compare survival in the two treatment groups using log-rank test

```{r}
# Log rank test
# To test if the is a difference between two survival curves

survdiff(Surv(time, event == 1) ~ tx, data = data.time)


```
This yields a low p-value, thus the HO cannot be accepted and the survival curves are significantly different from each other


Logistic Regression

```{r}
data <- read.delim("Logistic.txt",header = TRUE, sep = "", dec = ".")
data$AIDS_no=data$n-data$AIDS_yes
head(data,2) 

```




