#################
## Binary data ##
#################

Read the Logistic.txt into R
```{r}
data <- read.delim("Logistic.txt",header = TRUE, sep = "", dec = ".")
data$AIDS_no=data$n-data$AIDS_yes
head(data,2) 
```

Fit a logistic regression model

```{r}
df <- data.frame("AZT" = c(rep(1, 170), rep(0,168)), "AIDS" = c(rep(1, 25), rep(0, 145), rep(1, 44), rep(0, 124)))

head(df, )
         
length(AZT)
length(AIDS)

mod2 <- survreg(Surv(AIDS) ~ AZT, data = df, dist = "logistic")
summary(mod2)
         
```
Calculating the odds ratio

```{r}
exp(coef(mod2)[2]/mod2$scale)
```
Interpretation: The odds of survival until time t is 0.68 times greater for AZT = 1 than for AZT = 0

Confidence interval

```{r}
confint(mod2)

```
Test the hypothesis of no effect of AZT on AIDS

Likelihood ratio test

### The likelihood ratio test

The log likelihood ratio test is given as

$$log(LR) = log(L(\theta_{0}))-log(L(\hat \theta))$$

In this test $\theta_{0}$ is the MLE of no difference between the two treatment groups, meaning that it makes no difference whether a patient was treated with AZT. 

Likewise, $\hat \theta$ is the MLE of patients being treated with AZT. 

### Calculate thetas

First, $\theta_{0}$ is calculated. 

```{r}

#Formulating the likelihood function
binomial.log.likelihood = function(theta,success,size){
  -(sum(dbinom(x=success,size=size,prob=theta, log=TRUE)))}

theta.all = nlminb(start=0.5,objective=binomial.log.likelihood,lower=0,success=k,size=n)
theta.all$par
```

$\hat \theta$ is calculated. 

```{r}
theta_AZT_yes.opt=nlminb(start=0.5,binomial.log.likelihood,lower=0,success=data$AIDS_yes[1],size=data$n[1])

theta_AZT_yes.opt$par
```
LR is calculated. 

```{r}
(c <- theta.all$objective-theta_AZT_yes.opt$objective)
```

Given LR=c an approximate p-value can be calculated as

$$ P(\chi^2_{1} > -2 \cdot c)$$
```{r}
#(-2*c,df=1,lower.tail = F)

mod32 = glm(AIDS~AZT, family = binomial,data = df)

mod32$null.deviance-mod32$deviance
1-pchisq(mod32$null.deviance-mod32$deviance, df=1)
```
Low p-value meaning that AZT has an effect



The Wald test

The null hypothesis of log(scale) = 0 this can be seen directly from the R output
The null hypothesis is rejected on all levels as the p-value is 2e-16 



```{r}

```


#######################
## Suvival time data ##
#######################

Reading the data into R

```{r}
data.time <- read.delim("actg320.txt",header = TRUE, sep = "", dec = ".")
head(data.time,2) 
```

Number of patients getting AIDS or dying in the two groups
```{r}
# Treated group
(sum(data.time$event == 1 & data.time$tx == 1 ))


# Non treated group
(sum(data.time$event == 1 & data.time$tx == 0 ))

```
The total follow up time
follow up time = sum of the time that you follow all of the participants
```{r}
print('Follow up for new treatment group')
(sum(data.time$time[data.time$tx == 0])/365)


print('Follow up for new treatment group')
(sum(data.time$time[data.time$tx == 1])/365)
```
Plotting survival function in the two treatment groups

```{r}
library(survival)

Surv.ByTreat <- survfit(Surv(time, event == 1) ~ tx,conf.type = "log-log",data = data.time)

plot(Surv.ByTreat, conf.int = T, las = 1, xlab = "Days",ylab = "Estimated Survival Prob.", col=2:3, lwd = 2, ylim=c(0.8,1))
legend("bottomleft", col=2:3, c("Control","Treated"), lwd=2)


```
From the above plot it seems that the treated group is doing best

Plot the cumulative incidence functions

```{r}
plot(Surv.ByTreat, fun=function(x) { 1- x }, conf.int = F,las = 1, xlab = "Days",ylab = "Estimated Failure Prob.", col=2:3, lwd = 2)
legend("topleft", col=2:3, c("Control","Treated"), lwd=2)

# I would prefer the survival function as it is more positive to think about a survival probability than it is about a failure probability

```

Compare survival in the two treatment groups using log-rank test

```{r}
# Log rank test
# To test if the is a difference between two survival curves

survdiff(Surv(time, event == 1) ~ tx, data = data.time)


```
This yields a low p-value, thus the HO cannot be accepted and the survival curves are significantly different from each other


Logistic Regression

```{r}
data <- read.delim("Logistic.txt",header = TRUE, sep = "", dec = ".")
data$AIDS_no=data$n-data$AIDS_yes
head(data,2) 

```

################################
## Parametric survival models ##
################################

Fit parametric survival models containing treatment (tx) and CD4 count as exploratory variables

```{r}
mod1 <- survreg(Surv(time, event) ~ tx + cd4,
data = data.time, dist = "weibull")

summary(mod1)
```
It looks like that sigma factor is significant for the model, therefore, an exponential regression model would not be a better fit. 
sigma factor

```{r}

mod3 <- survreg(Surv(time, event) ~ cd4 + tx,
data = data.time, dist = "loglogistic")

summary(mod3)
```

AIC score for comparison

```{r}
c(AIC(mod1),AIC(mod3))
```
The loglogistic model is best because it has the lowest AIC score. 
Make a table of estimates for the parameters in the Loglogistic model including a 95% confidence interval

```{r}

confint(mod3)

```
Time Ratio log-logistic for th treatment effect and the effect of CD¤ count witg 50
```{r}
#Time ratio for bearings
exp(cbind(coef(mod3)[3],confint(mod3)[3, 1], confint(mod3)[3, 2]))
#Time ratio for loads
exp(cbind(coef(mod3)[2],confint(mod3)[2, 1], confint(mod3)[2, 2])*50)
```
```







