---
title: "R Notebook"
output: html_notebook
---

## Read the data tuno.txt into R.
The data is read, and the first few lines of the code is shown.
```{r}
data <- read.delim("tuno.txt",header = TRUE, sep = " ", dec = ".")
data$pow.obs.norm<-(data$pow.obs)/5000
head(data,6)
```


### 1. Formulate an initial model

The model that we consider is:
$$y=f(w_s)=\beta_0+\beta_1w_s+\beta_2w_s^2$$
Where y is the observed power production and f is a function of wind speed.
```{r}
mod1=lm(pow.obs.norm~ ws30+I(ws30^2), data=data)
summary(mod1)
AIC(mod1)
```
All parameters are significant.

```{r}
qqnorm(dnorm(data$pow.obs.norm, mean=mod1$fitted.values, sd=0.1648) )
qqline(dnorm(data$pow.obs.norm, mean=mod1$fitted.values, sd=0.1648))
```

### 2. You might consider non-normal models and/or normal model with data transformation. Further you might consider including wind direction. You should devellop a suited model for prediction of daily power prodiction. 

#### Models without tranformation 
```{r}
mod2=glm(pow.obs.norm~ ws30,family=Gamma,  data=data)
summary(mod2)
AIC(mod2)
```

```{r}
qqnorm(dnorm(data$pow.obs.norm, mean=mod2$fitted.values, sd=mod2$deviance) )
qqline(dnorm(data$pow.obs.norm, mean=mod2$fitted.values, sd=mod2$deviance))
```
```{r}
library(betareg)
mod8=betareg(pow.obs.norm~ ws30, data=data)
summary(mod8)
AIC(mod8)
```

```{r}
qqnorm(dnorm(data$pow.obs.norm, mean=mod8$fitted.values, sd=sqrt(sum((mod8$residuals)^2)/(length(data$pow.obs.norm)-2))))
qqline(dnorm(data$pow.obs.norm, mean=mod8$fitted.values, sd=sqrt(sum((mod8$residuals)^2)/(length(data$pow.obs.norm)-2))))
```


## Models using transformation 1, from project 1:
```{r, warning=FALSE}
trans1 <- function(lambda,y){
    y.l <- 1/lambda*log((y^lambda)/(1-y^lambda))
    return(y.l)}

## profile likelihood for lambda
lp.lambda1 <- function(lambda,y){
    n <- length(y)
    y.l <- trans1(lambda ,y)
    sigmasq <- 1/n * sum((y.l-mean(y.l))^2)
    -n/2 * log(sigmasq) + sum(log(abs(-1/(y*(-1 + y^lambda)))))}

(opt.lambda.trans1=optimize(lp.lambda1,c(0.000001,70),y=data$pow.obs.norm, maximum=TRUE))

```

```{r}
mod3=glm(trans1(opt.lambda.trans1$maximum,data$pow.obs.norm)~ ws30,family=gaussian,  data=data)
summary(mod3)
AIC(mod3)
```

```{r}
qqnorm(dnorm(trans1(opt.lambda.trans1$maximum,data$pow.obs.norm), mean=mod3$fitted.values, sd=mod3$deviance) ,main="Distribution after boxcox transformation")
qqline(dnorm(trans1(opt.lambda.trans1$maximum,data$pow.obs.norm), mean=mod3$fitted.values, sd=mod3$deviance))

```
## Models using transformation 2 from project 1:
```{r, warning=FALSE}
trans2 <- function(lambda,y){
    y.l <- 2*log((y^lambda)/(1-y)^(1-lambda))
    return(y.l)}

## profile likelihood for lambda
lp.lambda2 <- function(lambda,y){
    n <- length(y)
    y.l <- trans2(lambda ,y)
    sigmasq <- 1/n * sum((y.l-mean(y.l))^2)
    -n/2 * log(sigmasq) + sum(log(abs(((4*lambda - 2)*y - 2*lambda)/(y*(-1 + y)))))}
    
(opt.lambda.trans2=optimize(lp.lambda2,c(0.000001,1),y=data$pow.obs.norm, maximum=TRUE))

```
```{r}
mod7=glm(trans2(opt.lambda.trans2$maximum,data$pow.obs.norm)~ ws30,family=gaussian,  data=data)
summary(mod7)
AIC(mod7)
```
```{r}
qqnorm(dnorm(trans2(opt.lambda.trans2$maximum,data$pow.obs.norm), mean=mod7$fitted.values, sd=mod7$deviance))
qqline(dnorm(trans2(opt.lambda.trans2$maximum,data$pow.obs.norm), mean=mod7$fitted.values, sd=mod7$deviance))
```


## Models using Boxcox transformation: 
```{r}
## box-cox transformation
bc.trans <- function(lambda,y){
    y.l <- (y^lambda-1)/lambda
    
    if(lambda==0){y.l <- log(y)}
    return(y.l)}

## profile likelihood for lambda
lp.lambda <- function(lambda,y){
    n <- length(y)
    y.l <- bc.trans(lambda ,y)
    sigmasq <- 1/n * sum((y.l-mean(y.l))^2)
    -n/2 * log(sigmasq) + (lambda-1)*sum(log(y))}

(opt.lambda.boxcox=optimize(lp.lambda,c(-2,2),y=data$pow.obs.norm,maximum=TRUE))
```

```{r}
mod6=glm(bc.trans(opt.lambda.boxcox$maximum,data$pow.obs.norm)~ ws30,family=gaussian,  data=data)
summary(mod6)
AIC(mod6)
```

```{r}
qqnorm(dnorm(bc.trans(opt.lambda.boxcox$maximum,data$pow.obs.norm), mean=mod6$fitted.values, sd=mod6$deviance))
qqline(dnorm(bc.trans(opt.lambda.boxcox$maximum,data$pow.obs.norm), mean=mod6$fitted.values, sd=mod6$deviance))
```

