---
title: "Financial Part 3"
---

```{r}
data <- read.csv("finance_data.csv", header=TRUE,sep=";")
data$time=as.Date(data$time)
head(data,6)
```

```{r}
plot(data$time, data$SLV, type="l")
```

## Mixture models
### a) Fit normal mixture models with 2 and 3 components

Kig på Jans kode fra uge 10. Der er det implimenteret for en possion distribution og optimerer derfor over en parameter. Her skal det implementeres for en normal distribution og derfor for to parametre. 

```{r}


## Natural to working parameters
re.norm.mix.pn2pw <- function(m, mu, sigma, delta) {
  if(sum(delta) >= 1) {
    print("sum(delta) should be < 1")
    return()
  }
  t.sigma <- log(c(sigma[1],sigma[-1]-sigma[-m]))
  t.delta <- log(delta/(1 - sum(delta)))
  return(list(mu = mu, t.sigma = t.sigma, t.delta = t.delta))
}

## Working to natural parameters
re.norm.mix.pw2pn <- function(m, mu, t.sigma, t.delta){
  if(m == 1){
    return(exp(t.sigma))
  }
  sigma <- cumsum(exp(t.sigma))
  delta <- exp(t.delta)/(1 + sum(exp(t.delta)))
  delta <- c(1 - sum(delta),delta)
  return(list(mu = mu, sigma = sigma, delta = delta))
}


## Normal mixture: transform
## Natural to working parameters
norm.mix.pn2pw <- function(m, mu, sigma, delta) {
  if(sum(delta) >= 1) {
    print("sum(delta) should be < 1")
    return()
  }
  t.sigma <- log(sigma)
  t.delta <- log(delta/(1 - sum(delta)))
  return(list(mu = mu, t.sigma = t.sigma, t.delta = t.delta))
}

## Working to natural parameters
norm.mix.pw2pn <- function(m, mu, t.sigma, t.delta){
  if(m == 1){
    return(exp(t.sigma))
  }
  sigma <- exp(t.sigma)
  delta <- exp(t.delta)/(1 + sum(exp(t.delta)))
  delta <- c(1 - sum(delta),delta)
  return(list(mu = mu, sigma = sigma, delta = delta))
}

## Negative log likelihood
nll <- function(theta, m=2, x=data$SLV){
  if(m == 1) {
    return(-sum(pnorm(x, theta[1], exp(theta[2]), log=TRUE))) 
  }
  mu <- theta[1:m]
  t.sigma <- theta[(m+1):(2*m)]
  t.delta <- theta[(2*m+1):(3*m-1)]
  n.pars <- norm.mix.pw2pn(m, mu, t.sigma, t.delta)
  n <- length(x)
  nll <- 0
  for(i in 1:n) {
    nll <- nll - log(sum(n.pars$delta * dnorm(x[i], mu, n.pars$sigma)))
  }
  return(nll)
}

```


```{r}
## Estimation with 2 distributions
m <- 2; 

## Initial values
mu <- mean(data$SLV)*c(2,1/2)
sigma <-sd(data$SLV)*c(1,1)
delta <- c(0.01)

## Working parameters
wpars <- norm.mix.pn2pw(m, mu, sigma, delta)
theta <- c(wpars$mu, wpars$t.sigma, wpars$t.delta)

## MLE
opt2 <- nlminb(theta, nll, m = m, x = data$SLV)

## Natural parameters
npars2 <- norm.mix.pw2pn(m, opt2$par[1:m], opt2$par[(m+1):(2*m)], opt2$par[(2*m+1):(3*m-1)])

npars2 
```

```{r}
mix.dist <- function(x ,npars){
  sum(npars$delta * dnorm(x, mean = npars$mu, sd = npars$sigma))
}

## Plot
par(mfrow=c(1,1))
hist(data$SLV, prob=TRUE, nclass=60)
lines(seq(min(data$SLV), max(data$SLV), 0.001), sapply(seq(min(data$SLV), max(data$SLV), 0.001), mix.dist, npars=npars2), col=2)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars2$delta[1]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars2$mu[1], npars2$sigma[1]), col=3)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars2$delta[2]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars2$mu[2], npars2$sigma[2]), col=3)
legend("topleft",  c("2 components",'Components seperately'), col=c("red",'green'), lty=1, cex=0.7)

```


```{r}
## Estimation with 3 distributions
m <- 3;

## Initial values 
mu <- mean(data$SLV)*c(1/2,1,3/2)
sigma <- sd(data$SLV)*c(1/2,1,3/2);
delta <- c(1/3,1/3)

## Working parameters
wpars <- norm.mix.pn2pw(m, mu, sigma, delta)
theta <- c(wpars$mu, wpars$t.sigma, wpars$t.delta)

## MLE
opt3 <-nlminb(theta, nll, m = m, x = data$SLV)

## Natural parameters
npars3 <- norm.mix.pw2pn(m, opt3$par[1:m], opt3$par[(m+1):(2*m)], opt3$par[(2*m+1):(3*m-1)])

```


```{r}
mix.dist <- function(x ,npars){
  sum(npars$delta * dnorm(x, mean = npars$mu, sd = npars$sigma))
}

## Plot
par(mfrow=c(1,1))
hist(data$SLV, prob=TRUE, nclass=60)
lines(seq(min(data$SLV), max(data$SLV), 0.001), sapply(seq(min(data$SLV), max(data$SLV), 0.001), mix.dist, npars=npars3), col=2)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars3$delta[1]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars3$mu[1], npars3$sigma[1]), col=3)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars3$delta[2]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars3$mu[2], npars3$sigma[2]), col=3)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars3$delta[3]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars3$mu[3], npars3$sigma[3]), col=3)
legend("topleft",  c("3 components",'Components seperately'), col=c("red",'green'), lty=1, cex=0.7)


```
```{r}
## Model check
AIC <- 2*c(opt2$objective, opt3$objective) + 2*c(length(opt2$par), length(opt3$par))
AIC

## Deviance 
1-pchisq(-2*(opt3$objective-opt2$objective),df=length(opt3$par)-length(opt2$par))

```

Model 2 is slightly better than model 3. The  best model in financial part 1 was the t-distribution with an AIC score of -1492. Therefore this is the overall best model. 


### b) Report confidence interval for the parameters in the best mixture model. 

```{r}
library(numDeriv)
H <- hessian(nll, opt2$par)
se<- sqrt(diag(solve(H)))
(CI1 <- opt2$par[1]+c(-1,1)*se[1]*qnorm(0.975))
(CI2<- opt2$par[2]+c(-1,1)*se[1]*qnorm(0.975))
```
The mean value is centered around zero. Since, sigma is small, little variance in the data is expected. Better interpretation??

## c) Make a profile likelihood plot of one of the variance parameters in the two component model.  

```{r}
source("A1.R")
## Profile likelihood for sigma 1 given working parameters
lp.sigma1 <- function(sigma1, m, x, pars0){
  ## Fun for inner optim
  fun.tmp <- function(theta, sigma1, m, x){
    pars <- c(theta[1:m], log(sigma1), theta[-(1:m)])
    nll(pars, m, x)
  }
  nlminb(pars0, fun.tmp, sigma1 = sigma1, m = m, x = x)$objective    
}

## Estimation with 2 distributions
m <- 2; 

## Initial values
mu <- mean(data$SLV)*c(1/2,3/2)
sigma <-sd(data$SLV)*c(1/2,3/2)
delta <- c(0.1)

## Working parameters
wpars <- norm.mix.pn2pw(m, mu, sigma, delta)
theta0 <- c(wpars$mu, wpars$t.sigma, wpars$t.delta)
theta <- c(theta0[1:m],theta0[(m+2):(3*m-1)])

sigma1 <- seq(0.01, 0.1, length=100)

## profile likeihood
pnll <- sapply(sigma1, lp.sigma1, m = m, x = data$SLV, pars0 = theta)

## Plot the profile likelihood
plot(sigma1,exp(-(pnll-min(pnll))),type="l", ylim=c(0,1))
lines(range(sigma1),
      c(1,1)*exp(-qchisq(0.95,df=1)/2),col=2,lty=2,lwd=2)
rug(npars2$sigma,col=2,lwd=2)
```
We have two maxima in the profile likelihood. Meaning we don't know which one is better.


### d) Reparametrize

Hvordan fungerer denne her reparametrisering



```{r}
## Natural to working parameters
re.norm.mix.pn2pw <- function(m, mu, sigma, delta) {
  if(sum(delta) >= 1) {
    print("sum(delta) should be < 1")
    return()
  }
  t.sigma <- log(c(sigma[1],sigma[-1]-sigma[-m]))
  t.delta <- log(delta/(1 - sum(delta)))
  return(list(mu = mu, t.sigma = t.sigma, t.delta = t.delta))
}

## Working to natural parameters
re.norm.mix.pw2pn <- function(m, mu, t.sigma, t.delta){
  if(m == 1){
    return(exp(t.sigma))
  }
  sigma <- cumsum(exp(t.sigma))
  delta <- exp(t.delta)/(1 + sum(exp(t.delta)))
  delta <- c(1 - sum(delta),delta)
  return(list(mu = mu, sigma = sigma, delta = delta))
}

## Negative log likelihood
re.nll <- function(theta, m, x){
  if(m == 1) {
    return(-sum(pnorm(x, theta[1], exp(theta[2]), log=TRUE))) 
  }
  mu <- theta[1:m]
  t.sigma <- theta[(m+1):(2*m)]
  t.delta <- theta[(2*m+1):(3*m-1)]
  n.pars <- re.norm.mix.pw2pn(m, mu, t.sigma, t.delta)
  n <- length(x)
  nll <- 0
  for(i in 1:n) {
    nll <- nll - log(sum(n.pars$delta * dnorm(x[i], mu, n.pars$sigma)))
  }
  return(nll)
}

```


```{r}

## Profile likelihood for sigma 1 given working parameters
re.lp.sigma1 <- function(sigma1, m, x, pars0){
  ## Fun for inner optim
  fun.tmp <- function(theta, sigma1, m, x){
    pars <- c(theta[1:m], log(sigma1), theta[-(1:m)])
    re.nll(pars, m, x)
  }
  nlminb(pars0, fun.tmp, sigma1 = sigma1, m = m, x = x)$objective    
}

m <- 2; 

## Initial values
mu <- mean(data$SLV)*c(1/2,3/2)
sigma <-sd(data$SLV)*c(1/2,3/2)
delta <- c(0.1)

## Working parameters
wpars <- re.norm.mix.pn2pw(m, mu, sigma, delta)
theta0 <- c(wpars$mu, wpars$t.sigma, wpars$t.delta)
theta <- c(theta0[1:m],theta0[(m+2):(3*m-1)])

sigma1 <- seq(0.01, 0.05, length=100)

## profile likeihood
pnll <- sapply(sigma1, re.lp.sigma1, m = m, x = data$SLV, pars0 = theta)

## Plot the profile likelihood
plot(sigma1,exp(-(pnll-min(pnll))),type="l", ylim=c(0,1))
lines(range(sigma1),
      c(1,1)*exp(-qchisq(0.95,df=1)/2),col=2,lty=2,lwd=2)
rug(npars2$sigma[1],col=2,lwd=2)


```

### D. Discuss Interpretation of the models

We have a model that shows the distribution of the weekly return of the given ETF.
This is a two component mixture model meaning that is must consist of two underlying components. Where one component has significantly more influence than the other.

ANYTHING ELSE???


# Hidden Markov Models

### a) Fit two and three state normal Hidden Markov Models

```{r}
norm.HMM.pn2pw <- function(m, mu, sigma, gamma) {                                              
  t.sigma <- log(sigma)
  t.gamma <- NULL                              
  if (m > 1) {                                            
    foo     <- log(gamma / diag(gamma))           
    t.gamma <- as.vector(foo[!diag(m)])             
  }                                             
  parvect <- c(as.vector(mu), as.vector(t.sigma), 
               as.vector(t.gamma))
  parvect
}  

norm.HMM.pw2pn <- function(m, parvect) {
  mu    <- parvect[1:m]
  epar  <- exp(parvect[-(1:m)]) 
  sigma <- epar[1:m]
  gamma <- diag(m)                                    
  if (m > 1) {                                                  
    gamma[!gamma] <- epar[(m+1):(m*m)]
    gamma         <- gamma / apply(gamma, 1, sum)          
  }                                                   
  delta <- solve(t(diag(m) - gamma + 1), rep(1, m))          
  list(mu = mu, sigma = sigma, gamma = gamma, delta = delta)           
}  

norm.HMM.mllk <- function(parvect, x, m, ...) {
  n        <- length(x)                            
  pn       <- norm.HMM.pw2pn(m, parvect)            
  P        <- rep(NA, m)
  lscale   <- 0                                    
  foo      <- pn$delta  
  for (i in 1:n) {
    for (j in 1:m) {
      P[j] <- dnorm(x[i], mean = pn$mu[j], sd = pn$sigma[j])
    }
    foo    <- foo %*% pn$gamma * P
    sumfoo <- sum(foo)
    lscale <- lscale + log(sumfoo)
    foo    <- foo / sumfoo
  }
  mllk     <- -lscale
  mllk
}

norm.HMM.mle <- function(x, m, mu0, sigma0, gamma0,...) {                                                      
  parvect0 <- norm.HMM.pn2pw(m, mu0, sigma0, gamma0)         
  mod      <- nlm(norm.HMM.mllk, parvect0, x = x, m = m) 
  pn       <- norm.HMM.pw2pn(m, mod$estimate)            
  mllk     <- mod$minimum                              
  np       <- length(parvect0)
  AIC      <- 2 * (mllk + np)                              
  n        <- sum(!is.na(x))                            
  BIC      <- 2 * mllk + np * log(n)                         
  list(mu = pn$mu, sigma = pn$sigma, gamma = pn$gamma, delta = pn$delta, 
       code = mod$code, mllk = mllk, AIC = AIC, BIC = BIC)   
}

########################################################################################################################

## 2 - state 
## Initial values
m <- 2
mu0 <- mean(data$SLV)*c(1/2,3/2)
sigma0 <- sd(data$SLV)*c(1/2,3/2)
gamma0 <- matrix(0.01,ncol = m, nrow = m)
diag(gamma0) <- 1-(m - 1)*gamma0[1,1]

## optimize
fit2 <- norm.HMM.mle(data$SLV, m, mu0, sigma0, gamma0)
fit2


## 3 - state 
## Initial values
m <- 3
mu0 <- mean(data$SLV)*c(1/2,1,3/2)
sigma0 <- sd(data$SLV)*c(1/2,1,3/2)
gamma0 <- matrix(0.01,ncol = m, nrow = m)
diag(gamma0) <- 1-(m - 1)*gamma0[1,1]

## optimize
fit3 <- norm.HMM.mle(data$SLV, m, mu0, sigma0, gamma0)
fit3
```








