---
title: "Financial Part 3"
---

```{r}
data <- read.csv("finance_data.csv", header=TRUE,sep=";")
data$time=as.Date(data$time)
head(data,6)
```
## Mixture models
### a) Fit normal mixture models with 2 and 3 components
```{r}
library(MASS)
library(mclust, quietly=TRUE)

fit3 = Mclust(data$SLV, G=3, model="V")
summary(fit3)

```

```{r}
fit2 = Mclust(data$SLV, G=2, model="V")
summary(fit2)
```
The normal mixture model with 3 components is better than the model with 2 components, looking at both the log-likelihood and the BIC score. 

### b) Report confidence intervals for the parameters and give an interpretation of these intervals. 
```{r}
fit3$parameters$mean[1]+c(1,-1)*sqrt(fit3$parameters$variance$sigmasq[1])*qnorm(0.975)
fit3$parameters$mean[2]+c(1,-1)*sqrt(fit3$parameters$variance$sigmasq[2])*qnorm(0.975)
fit3$parameters$mean[3]+c(1,-1)*sqrt(fit3$parameters$variance$sigmasq[3])*qnorm(0.975)

```
All the confidence intervals for the mean parameters include zero, meaning that all the gaussian distributions are centered around zero. That probably okay, given that this is a confidence interval for the mean. 

## c) Make a profile likelihood plot of one of the variance parameters in the two component model.  

```{r}

```


