---
title: "Financial Part 3"
---

```{r}
data <- read.csv("finance_data.csv", header=TRUE,sep=";")
data$time=as.Date(data$time)
head(data,6)
```

```{r}
plot(data$time, data$SLV, type="l")
```

## Mixture models
### a) Fit normal mixture models with 2 and 3 components

Kig på Jans kode fra uge 10. Der er det implimenteret for en possion distribution og optimerer derfor over en parameter. Her skal det implementeres for en normal distribution og derfor for to parametre. 

```{r}
## Normal mixture: transform
## Natural to working parameters
norm.mix.pn2pw <- function(m, mu, sigma, delta) {
  if(sum(delta) >= 1) {
    print("sum(delta) should be < 1")
    return()
  }
  t.sigma <- log(sigma)
  t.delta <- log(delta/(1 - sum(delta)))
  return(list(mu = mu, t.sigma = t.sigma, t.delta = t.delta))
}

## Working to natural parameters
norm.mix.pw2pn <- function(m, mu, t.sigma, t.delta){
  if(m == 1){
    return(exp(t.sigma))
  }
  sigma <- exp(t.sigma)
  delta <- exp(t.delta)/(1 + sum(exp(t.delta)))
  delta <- c(1 - sum(delta),delta)
  return(list(mu = mu, sigma = sigma, delta = delta))
}

## Negative log likelihood
nll <- function(theta, m=2, x=data$SLV){
  if(m == 1) {
    return(-sum(pnorm(x, theta[1], exp(theta[2]), log=TRUE))) 
  }
  mu <- theta[1:m]
  t.sigma <- theta[(m+1):(2*m)]
  t.delta <- theta[(2*m+1):(3*m-1)]
  n.pars <- norm.mix.pw2pn(m, mu, t.sigma, t.delta)
  n <- length(x)
  nll <- 0
  for(i in 1:n) {
    nll <- nll - log(sum(n.pars$delta * dnorm(x[i], mu, n.pars$sigma)))
  }
  return(nll)
}

```


```{r}
## Estimation with 2 distributions
m <- 2; 

## Initial values
mu <- mean(data$SLV)*c(1/2,3/2)
sigma <-sd(data$SLV)*c(1/2,3/2)
delta <- c(1/2)

## Working parameters
wpars <- norm.mix.pn2pw(m, mu, sigma, delta)
theta <- c(wpars$mu, wpars$t.sigma, wpars$t.delta)

## MLE
opt2 <- nlminb(theta, nll, m = m, x = data$SLV)

## Natural parameters
npars2 <- norm.mix.pw2pn(m, opt2$par[1:m], opt2$par[(m+1):(2*m)], opt2$par[(2*m+1):(3*m-1)])

npars2 
```

```{r}
mix.dist <- function(x ,npars){
  sum(npars$delta * dnorm(x, mean = npars$mu, sd = npars$sigma))
}

## Plot
par(mfrow=c(1,1))
hist(data$SLV, prob=TRUE, nclass=60)
lines(seq(min(data$SLV), max(data$SLV), 0.001), sapply(seq(min(data$SLV), max(data$SLV), 0.001), mix.dist, npars=npars2), col=2)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars2$delta[1]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars2$mu[1], npars2$sigma[1]), col=3)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars2$delta[2]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars2$mu[2], npars2$sigma[2]), col=3)
legend("topleft",  c("2 components",'Components seperately'), col=c("red",'green'), lty=1, cex=0.7)

```


```{r}
## Estimation with 3 distributions
m <- 3;

## Initial values 
mu <- mean(data$SLV)*c(1/2,1,3/2)
sigma <- sd(data$SLV)*c(1/2,1,3/2);
delta <- c(1/3,1/3)

## Working parameters
wpars <- norm.mix.pn2pw(m, mu, sigma, delta)
theta <- c(wpars$mu, wpars$t.sigma, wpars$t.delta)

## MLE
opt3 <-nlminb(theta, nll, m = m, x = data$SLV)

## Natural parameters
npars3 <- norm.mix.pw2pn(m, opt3$par[1:m], opt3$par[(m+1):(2*m)], opt3$par[(2*m+1):(3*m-1)])

```


```{r}
mix.dist <- function(x ,npars){
  sum(npars$delta * dnorm(x, mean = npars$mu, sd = npars$sigma))
}

## Plot
par(mfrow=c(1,1))
hist(data$SLV, prob=TRUE, nclass=60)
lines(seq(min(data$SLV), max(data$SLV), 0.001), sapply(seq(min(data$SLV), max(data$SLV), 0.001), mix.dist, npars=npars3), col=2)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars3$delta[1]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars3$mu[1], npars3$sigma[1]), col=3)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars3$delta[2]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars3$mu[2], npars3$sigma[2]), col=3)
lines(seq(min(data$SLV), max(data$SLV), 0.001), npars3$delta[3]*dnorm(seq(min(data$SLV), max(data$SLV), 0.001), npars3$mu[3], npars3$sigma[3]), col=3)
legend("topleft",  c("2 components",'Components seperately'), col=c("red",'green'), lty=1, cex=0.7)


```
```{r}
## Model check
AIC <- 2*c(opt2$objective, opt3$objective) + 2*c(length(opt2$par), length(opt3$par))
AIC

## Deviance 
1-pchisq(-2*(opt3$objective-opt2$objective),df=length(opt3$par)-length(opt2$par))

```

Model 2 is slightly better than model 3. The  best model in financial part 1 was the t-distribution with an AIC score of -1492. Therefore this is the overall best model. 


### b) Report confidence interval for the parameters in the best mixture model. 

```{r}
library(numDeriv)
H <- hessian(nll, opt2$par)
se<- sqrt(diag(solve(H)))
(CI1 <- opt2$par[1]+c(-1,1)*se[1]*qnorm(0.975))
(CI2<- opt2$par[2]+c(-1,1)*se[1]*qnorm(0.975))
```
The mean value is centered around zero. Since, sigma is small, little variance in the data is expected. 

## c) Make a profile likelihood plot of one of the variance parameters in the two component model.  

```{r}
## Profile likelihood for sigma 1 given working parameters
lp.sigma1 <- function(sigma1, m, x, pars0){
  ## Fun for inner optim
  fun.tmp <- function(theta, sigma1, m, x){
    pars <- c(theta[1:m], log(sigma1), theta[-(1:m)])
    nll(pars, m, x)
  }
  nlminb(pars0, fun.tmp, sigma1 = sigma1, m = m, x = x)$objective    
}

```




## Markov Chain
```{r}
y <- data$SLV

```

```{r}
source("A1.R")
y <- data$SLV

## 2 - state 
## Initial values
m <- 2
lambda0 <- quantile(y,c(0.25,0.75))
gamma0 <- matrix(0.05,ncol=m,nrow=m)
diag(gamma0) <- 1-(m-1)*gamma0[1,1]


## optimize
fit2 <- pois.HMM.mle(y,m,lambda0,gamma0)
fit2

```

```{r}
## working parameters
parvect  <- pois.HMM.pn2pw(m,fit3$lambda,fit3$gamma)
## Optimize (hessian = TRUE return hessian)
mod <- nlm(pois.HMM.mllk,parvect,x=y,m=m,
            hessian=TRUE)  
mod

## Organize the result
parvect <- mod$estimate
names(parvect) <- c("lambda1","lambda2","lambda3","tau21",
                    "tau31","tau12","tau32","tau13","tau23")

se <- sqrt(diag(solve(mod$hessian)))

## Working pars + standard error
round(cbind(parvect,se),digits=2) ## note se of tau31
fit3$gamma
```

