---
title: "R Notebook"
output: html_notebook
---

## Read the data tuno.txt into R.
The data is read, and the first few lines of the code is shown.
```{r}
data <- read.delim("tuno.txt",header = TRUE, sep = " ", dec = ".")
data$pow.obs.norm<-(data$pow.obs)/5000
head(data,6)
```


### 1. Formulate an initial model

The model that we consider is:
$$y=f(w_s)=\beta_0+\beta_1w_s+\beta_2w_s^2$$
Where y is the observed power production and f is a function of wind speed.
```{r}
mod1=lm(pow.obs.norm~ ws30+I(ws30^2), data=data)
summary(mod1)
AIC(mod1)
```
All parameters are significant.

```{r}
library(car)
qqPlot(mod1$residuals)
```

### 2. You might consider non-normal models and/or normal model with data transformation. Further you might consider including wind direction. You should devellop a suited model for prediction of daily power produAction. 

#### Models without tranformation 
```{r}
mod2=glm(pow.obs.norm~ ws30,family=Gamma,  data=data)
summary(mod2)
AIC(mod2)
```

```{r}
qqPlot(mod2$residuals, main = "qq-plot gamma regression")
```
```{r}
library(betareg)
mod8=betareg(pow.obs.norm~ ws30, data=data)
summary(mod8)
AIC(mod8)
```

```{r}
qqPlot(mod8$residuals, main = "qq-plot beta regression")
```


## Models using transformation 1, from project 1:
Optimer lambda således at residualerne fra modellen best fitter en normalfordeling. 


#### Wind Speed and Wind Speed Squared
```{r, warning=FALSE}
trans1 <- function(lambda,y){
    y.l <- 1/lambda*log((y^lambda)/(1-y^lambda))
    return(y.l)}

## profile likelihood for lambda
lp.lambda1 <- function(lambda,y){
    mod <- lm(trans1(lambda,y)~data$ws30+I(data$ws30^2))
    length(y)/2*log(summary(mod)$sigma^2) - sum(log(abs(1/(y*(1 - y^lambda)))))
    }

#plot(seq(0.01,1,0.01), sapply(seq(0.01,1,0.01), lp.lambda1, y=data$pow.obs.norm))

(opt.lambda.trans1=optimize(lp.lambda1,c(0,1),y=data$pow.obs.norm))

```
```{r}
mod3=glm(trans1(opt.lambda.trans1$minimum,data$pow.obs.norm)~ ws30+I(data$ws30^2),family=gaussian,  data=data)
summary(mod3)
AIC(mod3)
qqPlot(mod3$residuals)
```
We have tried to add a third order term, but it did not improve the model as it was not significant.

#### Wind speed, Wind speed squared and cosine of wind direction
```{r}
#Forsøg med sin og cos til wd

trans1 <- function(lambda,y){
    y.l <- 1/lambda*log((y^lambda)/(1-y^lambda))
    return(y.l)}

## profile likelihood for lambda
lp.lambda1 <- function(lambda,y){
    mod <- lm(trans1(lambda,y)~data$ws30+I(data$ws30^2)+cos(data$wd30))
    length(y)/2*log(summary(mod)$sigma^2) - sum(log(abs(1/(y*(1 - y^lambda)))))
    }

(opt.lambda.trans1=optimize(lp.lambda1,c(0,1),y=data$pow.obs.norm))

```
```{r}
library(car)
mod5=glm(trans1(opt.lambda.trans1$minimum,data$pow.obs.norm)~ data$ws30+I(data$ws30^2)+cos(data$wd30),family=gaussian,  data=data)
summary(mod5)
AIC(mod5)
qqPlot(mod5$residuals)
```

#### Wind speed, Wind speed squared and cosine of wind direction
```{r}
#Forsøg med sin og cos til wd

trans1 <- function(lambda,y){
    y.l <- 1/lambda*log((y^lambda)/(1-y^lambda))
    return(y.l)}

## profile likelihood for lambda
lp.lambda1 <- function(lambda,y){
    mod <- lm(trans1(lambda,y)~data$ws30+I(data$ws30^2)+cos(data$wd30))
    length(y)/2*log(summary(mod)$sigma^2) - sum(log(abs(1/(y*(1 - y^lambda)))))
    }

(opt.lambda.trans1=optimize(lp.lambda1,c(0,1),y=data$pow.obs.norm))

```

```{r}
library(car)
mod6=glm(trans1(opt.lambda.trans1$minimum,data$pow.obs.norm)~data$ws30+I(data$ws30^2)+cos(data$wd30),family=gaussian,  data=data)
summary(mod6)
AIC(mod6)
qqPlot(mod6$residuals)
plot(data$ws30,mod6$residuals)
```


### 3. Present the parameters of the final model, this include quantification of the uncertainty of parameters.
```{r}
Confint(mod6)
```

### 4. Give an interpretation of the parameters in the particular this should include presentation of any nonlinear functions (series expansions) of the exploratory variables.
Walt test on the parameters.

Hvad gør de forskellige værdier for vore power production
fx sin kan give noget svingning.

**cos(data$wd30)**
Leddet påvirker power production negativt, derfor har den størst positivt bidrag når den er lig -1.  Cosinus til $\pi$ er lig -1, det vil sige, at vi får mest ud af wind direction, når vinden kommer med en vinkel på 180 grader.

**data$ws30**
Wind speed påvirker production, når den has sin største værdi, ergo må power production være højest på meget blæsende dage. Påvirkning fra **data$ws30** vil være større end påvirkning fra **I(data$ws30^2)** så længe at wind speed ikke overskrider 54.5 (se plot nedenfor).Dette er også højest usandsynligt. 


```{r}
ws=0:100
plot(ws, 1.63430162*ws+(-0.03013925)*ws^2, type="l", ylim=c(-10,25))
abline(h=0)
```




### 5. Present the final model e.g. some graphical presentation of predictions under different scenarios of wind speed and wind direction. 
```{r}
# Function to plot the prediction result of regression model based on wind speed.
predictRegression.windSpeed <- function(x1.pred, x2, y, modelRegression, name){
      dataFramePredict <- data.frame(x1 = x1.pred, x2 = rep(mean(x2), length(x1.pred)))
      # Calculate predictions and intervals
      valuePred <- predict.lm(modelRegression, dataFramePredict, se.fit = TRUE)
      intervalPred <- predict.lm(modelRegression, newdata = dataFramePredict, interval = "pred", level = 0.95)
      intervalConf <- predict.lm(modelRegression, newdata = dataFramePredict, interval = "conf", level = 0.95)
     # plot
      plot(dataFramePredict$x1, valuePred$fit, pch = 20, ylim = c(min(y, intervalPred), max(y, intervalPred)), xlim = c(min(x1, dataFramePredict$x1), max(x1, dataFramePredict$x1)), xlab = name, ylab = 'Transformed, Normalized Power Production')
      matlines(dataFramePredict$x1, intervalPred, lty = c(1, 2, 2), col = c("black", "blue", "blue"), lwd = 2)
      points(x1, y)  # Transformed observations
      lines(dataFramePredict$x1, intervalConf[, 2], lty = 3, col = "red", lwd = 2)
      lines(dataFramePredict$x1, intervalConf[, 3], lty = 3, col = "red", lwd = 2)
      legend("bottomright", inset = .02, legend = c("Regression Curve", "Prediction Interval", "Confidence Intercal"), col = c("black", "blue", "red"), lty = c(1, 2, 3), lwd = c(2, 2, 2), cex = 0.8)
}
```

```{r}
x1.pred <- seq(0, 30, 0.1)
x1 = data$ws30
x2 =data$wd30
y = trans1(opt.lambda.trans1$minimum,data$pow.obs.norm)
fittrans1 = glm(y ~ 1 + x1 + I(x1^2) + cos(x2))
predictRegression.windSpeed(x1.pred, x2, y, fittrans1, "Wind Speed")
```
Sort i¨linje: regressionen
Rød linje: confidensinterval for regressionen
Blå linje: Interval hvori vi forventer værdier


```{r}
x1.pred <- seq(0, 2*3.1415, 0.1)
x2 = data$ws30
x1 =data$wd30
y = trans1(opt.lambda.trans1$minimum,data$pow.obs.norm)
fittrans1 = glm(y ~ 1 + x2 + I(x2^2) + cos(x1))
predictRegression.windSpeed(x1.pred, x2, y, fittrans1,"Wind Direction")
```

```{r}
plot(trans1(opt.lambda.trans1$minimum,data$pow.obs.norm), ylab = "Transformed, Normalized Power Production")
points(mod6$fitted.values, pch=16, col="Red")   

#lines(interval_l, col)
```

Holder wind direction mod trans data sæt en værdi ind for wind direction.

```{r}
plot(trans1(opt.lambda.trans1$minimum,data$pow.obs.norm))
```

