# Analysis of the binary data

### Read data

```{r}
data <- read.delim("Logistic.txt",header = TRUE, sep = "", dec = ".")
data$AIDS_no=data$n-data$AIDS_yes
head(data,2)

n <- sum(data[,3]) #Number of patient #n 
k <- sum(data[,2]) #Number of patients with AIDS #Sucesses #k
theta <- k/n #Probability that you have Aids
```

### Fit a logistic regression model for the binary outcome AIDS=”yes” versus AIDS=”no” with the explanatory variable treatment with AZT (Yes,NO).

```{r}
AZT = c(rep(1,170),rep(0,168))
AIDS = c(rep(1,25),rep(0,145),rep(1,44),rep(0,124))

#Creating a dataframe with data
df = data.frame(AZT,AIDS)
```

```{r}
#Logistic regression
mod1=glm(AIDS~AZT, family = binomial,data = df)
summary(mod1)
```

###  Present the odds ratio for the effect of AZT on AIDS with 95% confidence interval and interpret the result in words.

The odds ratio is defined as: 

$$
O R=\frac{D_{E} H_{N}}{D_{N} H_{E}}
$$

$$
\begin{array}{|r|cc|}
\hline & \text { Diseased } & \text { Healthy } \\
\hline \text { Exposed } & D_{E} & H_{E} \\
\text { Not exposed } & D_{N} & H_{N} \\
\hline
\end{array}
$$

```{r}
a = data$AIDS_yes[1]
b = data$AIDS_no[1]
c = data$AIDS_yes[2]
d = data$AIDS_no[2]

(OR=(a*d)/(b*c))

```
The OR is lower than one, meaning that probability of getting aids is lower when a patient is treated with AZT. 

```{r}
#with confidence interval #95% confidence interval
(lowerOR=exp(log(OR)-qnorm(0.975)*sqrt(1/a+1/b+1/c+1/d)))
(upperOR=exp(log(OR)+qnorm(0.975)*sqrt(1/a+1/b+1/c+1/d)))
```


The confidence interval does not include one, meaning that probability of getting aids is lower when a patient is treated with AZT. 

### Test the hypothesis of no effect of AZT on AIDS using:

#### The likelihood ratio test

The log likelihood ratio test is given as

$$log(LR) = log(L(\theta_{0}))-log(L(\hat \theta))=log(c)$$

Here $\theta_{0}$ is deviance of the model with only the intercept and $ \hat\theta$ is the deviance of the model including all parameters. 

Using the c value from above, an approximated p-value can be obtained by:
$$P(\chi^2_1>-2\cdot log(c))$$
!!!!!!!!!!!!!!!Hvorfor forsvinder 2-tallet????????????????

```{r}
#mod1 = glm(AIDS~AZT, family = binomial,data = df)

c = mod1$null.deviance-mod1$deviance
1-pchisq(c, df=1)
```
The null-hypothesis is rejected. There is a significan difference between the two models, meaning that the model with all parameters are significantly better. 


#### Wald test

Find the Wald confidence intervals. 

```{r}
confint(mod1)
```
In the confidence interval for AZT, zero is not included. Therefore, the AZT parameter is significant for the model. 

The Wald statistics for the AZT effect:

The null-hypothesis is the that theta is equal to 0. 

```{r}
z=coef(mod1)[2]/summary(mod1)$coefficients[2,2]
```
Based on Wald test statistics a p-value is calculated. 

```{r}
2*pnorm(z)
1-pchisq(z^2, df=1)
```
```


# Analysis of the survival data

## 1 Analyses of the survival time data

### Load data
```{r}
#Load data
data.time <- read.delim("actg320.txt",header = TRUE, sep = "", dec = ".")
head(data.time,2) 
```

### How many got AIDS in the two treament groups? 
```{r}

by(data.time$event, data.time$tx, sum)

```

### How long is the follow-up time?

```{r}
#From days to years
data.time$YearTime <- data.time$time/365.25

by(data.time$YearTime, data.time$tx, sum)

```
The follow-up time is a little longer for patients with treatment. 

### Plot the survival function in the two treatment groups, which groups seems to be doing best?
```{r}
library(survival)

Surv.Bytx <- survfit(Surv(time, event == 1) ~ tx,
conf.type = "log-log",data = data.time)

(Surv.Bytx)
```

```{r}
plot(Surv.Bytx, conf.int = T, las = 1, xlab = "Time",ylab = "Estimated Survival Prob.", col=c(2,3), lwd = 1)

legend("bottomleft", col=2:3, c("Not treated","Treated"), lwd=2)
```
The functions differ towards the end. It is seen that the probability of surviving is higher when a patient is treated. 

### Plot the cumulative incidence function for the two groups which do you prefer?

```{r}
plot(Surv.Bytx, fun=function(x) { 1- x }, conf.int = T,las = 1, xlab = "Time",
ylab = "Estimated Failure Prob.", col=c(2,3))
legend("topleft", col=c(2,3), c("Male","Female"))
```
The cumulative incidence function is best, since it is easier to see what is happening on this plot given than the incidence rate is fairly low.  

### Compare the survival in the two treatment groups using the Log-rank test

```{r}
survdiff(Surv(YearTime, event == 1) ~ tx, data = data.time)
```
The p-value is significant. The null-hypothesis is rejected, meaning that it does make a difference to receive treatment. 


## 1.2 Parametric survival models

## Fit parametric survival models containing treatment (tx) and CD4 count as exploratory variables

### Weibull model 
```{r}
mod1 <- survreg(Surv(time, event) ~ tx + cd4,
data = data.time, dist = "weibull")

summary(mod1)
```
It looks like that sigma factor is significant for the model, therefore, an exponential regression model would not be a better fit. 


### Log-logistic model

```{r}
mod3 <- survreg(Surv(time, event) ~ cd4 + tx,
data = data.time, dist = "loglogistic")

summary(mod3)
```
All parameters are significant.


### AIC 

To compare the 2 models the AIC score is calculated
```{r}
c(AIC(mod1),AIC(mod3))
```
The loglogistic model is best because it has the lowest AIC score. 

### Make a table of estimates for the parameters in the Loglogistic model including a 95% confidence interval

```{r}

confint(mod3)

```

### Time Ratio log-logistic for th treatment effect and the effect of CD¤ count witg 50

```{r}
#Time ratio for bearings
exp(cbind(coef(mod3)[3],confint(mod3)[3, 1], confint(mod3)[3, 2]))
#Time ratio for loads
exp(cbind(coef(mod3)[2],confint(mod3)[2, 1], confint(mod3)[2, 2])*50)
```
The survival time is increased with 132% when a patient is treated. 

The survival time is increased 183% for every 50 cd4 cells/milliliter. 

### Asses the goodness of fit of the loglogistic model using the Cox Snell residuals

```{r}
data.time$z <- (log(data.time$time)-mod3$linear.predictors)/mod3$scale
data.time$CS3 <- log(1+exp(data.time$z))
surv3 <- survfit(Surv(CS3, event==1)~1 , data = data.time)
plot(surv3$time, -log(surv3$surv))
abline(a=0, b=1, lty=2)
```
The fit looks okay. The models best in the early survival times. The longer time that passes, the worse it fits. 

### Give a grafical representation of model

```{r}
xrange <- range(data.time$time)
t <- seq(xrange[1],xrange[2],length=100)

#LOG LOGISTIC
coef3 <- mod3$coefficients

z31 <- (log(t)-(coef3[1]+coef3[2]+coef3[3]))/mod3$scale
z30 <- (log(t)-(coef3[1]+coef3[2]))/mod3$scale
S31 <- (1+exp(z31))^-1
S30 <- (1+exp(z30))^-1


# get the range for the x and y axis
xrange <- range(data.time$time)
yrange <- range(S30)


# set up the plot
plot(xrange, yrange, type="n", xlab="Hours since test start",
ylab="Probability of survival (cd4=1)",las=1)
#PLOT THE SURVIVAL FUNCTIONS
lines(t, S31, type="l", col=2, lty=1, lwd=2)
lines(t, S30, type="l", col=3, lty=1, lwd=2)
legend('topright',lwd=2,col=c(2,3),legend=c('LL with tx', 'LL without tx'))

```
It is seen that patients in general have higher probability of surviving if they are treated.  




