---
title: Survival
output: html_notebook
---

# Project 2 
## Analysis of the binary data
### Read the data into R:
```{r}
data <- read.delim("Logistic.txt",header = TRUE, sep = "", dec = ".")
data$AIDS_no=data$n-data$AIDS_yes
data
```

### Fit a logistic regression model:
```{r}
# AZT treatment =1, AZT no treatment=0:
AZT=c(rep(1,170),rep(0,168))
# AIDS =1, no AIDS=0
AIDS=c(rep(1,25),rep(0,(170-25)),rep(1,44),rep(0,(168-44)))
# Combine it to a dataframe
df = data.frame(AZT, AIDS) 
```

```{r}
# Performing a logistic regression
fit.glm=glm(AIDS~AZT,family=binomial, data=df)
fit.glm
```

The odds ratio: 
$$OR=\frac{a*d}{b*c}$$
$D_E$ - exposed and diseased
$H_N$ - Not exposed and healthy

```{r}
a=data$AIDS_yes[1]
b=data$AIDS_no[1]
c=data$AIDS_yes[2]
d=data$AIDS_no[2]
(OR=a*d/(b*c))
```
OR<1 means that if the patient is treated the probability of getting AIDS is lower.

Calculate a 95% confidence interval:
```{r}
(upperOR=exp(log(OR)+qnorm(0.975)*sqrt(1/a+1/b+1/c+1/d)))
(lowerOR=exp(log(OR)-qnorm(0.975)*sqrt(1/a+1/b+1/c+1/d)))
```
The 95% confidence interval does not include 1, meaning that is a patient is treated with AZT the probability of getting AIDS is lower.

## Analysis of the survival time data

## Descriptive statistics:

### Read the data into R:
```{r}
#Load data
data.time <- read.delim("actg320.txt",header = TRUE, sep = "", dec = ".")
head(data.time,5) 
```

### How many patients got AIDS or died in the two treatment groups?
```{r}
# Control treatment got AIDS or died
(sum(data.time$tx == 0 & data.time$event == 1))
# New treatment got AIDS or died
(sum(data.time$tx == 1 & data.time$event == 1))
```

### How long was the total follow-up time in the two groups?
```{r}
# Convert days to years
data.time$timeyear=data.time$time/365.25
# Calculate followup time for the two groups
by(data.time$timeyear,data.time$tx, sum)
```

We have longer followup time for treatment group 1 (new treatment) than 0 (control treatment).

### Plot the survival functions in the two treatment groups, which group seems to be doing best?
```{r}
library(survival)
Surv.bytx = survfit(Surv(timeyear,event==1)~tx, conf.type="log-log", data=data.time)
Surv.bytx
```

```{r}
plot(Surv.bytx, conf.int=T, las=1, xlab="Years Since Admission", ylab="Estimated Survival Prob", col=2:3, lwd=2, ylim=c(0.75,1))
legend("bottomleft",col=2:3, c("Control Treatment", "New Treatment"),lwd=2)
```
The probability of surviving is higher whit the new treatment.

### Plot the cumulative incidence functions for the two groups, which plot would you prefer?
```{r}
plot(Surv.bytx, fun=function(x) { 1- x }, conf.int = F, las = 1, xlab = "Years since admission",
ylab = "Estimated Failure Prob.", col=2:3, lwd = 2)
legend("topleft", col=2:3, c("Control Treatment", "New Treatment"), lwd=2)
```
I prefer the survival function, as it is more postive!

### Compare the survival in the two treatment groups using a log-rank test.
```{r}
survdiff(Surv(timeyear, event==1)~tx, data=data.time)
```

The p-value is 0.001 which means that there is a significant difference between the two groups.
