################
## Wind Power ##
################

Reading the data

```{r}
data <- read.delim("tuno.txt",header = TRUE, sep = " ", dec = ".")
data$pow.obs.norm<-(data$pow.obs)/5000
head(data,6)

```

Formulate an initial model f = b_0 + b_1*w_s + b_2*w_s^2

```{r}

mod1=lm(pow.obs.norm~ ws30+I(ws30^2), data=data)

summary(mod1)

AIC(mod1)
```
All parameters are significant

###############################################################
## Develop a suited model for prediction of power production ##
###############################################################

We also tried to add wind direction

```{r}
mod4=lm(pow.obs.norm~ ws30+wd30+I(ws30^2)+I(wd30^2), data=data)

summary(mod4)

AIC(mod4)
```

Wind direction is also slightly significant
It does not improve the model much
Using the generalized linear model with a Gamma link function appears to be better

```{r}
#This is the best model so far
mod2=glm(pow.obs.norm~ ws30,family=Gamma,  data=data)
summary(mod2)
AIC(mod2)
```

Using transformation 1 from project

```{r}
trans1 <- function(lambda,y){
    y.l <- 1/lambda*log((y^lambda)/(1-y^lambda))
    return(y.l)}

## profile likelihood for lambda
lp.lambda1 <- function(lambda,y){
    n <- length(y)
    y.l <- trans1(lambda ,y)
    sigmasq <- 1/n * sum((y.l-mean(y.l))^2)
    -n/2 * log(sigmasq) + sum(log(abs(-1/(y*(-1 + y^lambda)))))}

(opt.lambda.trans1=optimize(lp.lambda1,c(0.000001,70),y=data$pow.obs.norm, maximum=TRUE))
```
```{r}
mod3=glm(trans1(opt.lambda.trans1$maximum,data$pow.obs.norm)~ ws30,family=gaussian,  data=data)

summary(mod3)

AIC(mod3)
```

```{r}
mod3=glm(trans1(opt.lambda.trans1$maximum,data$pow.obs.norm)~ ws30,family = gaussian,  data=data)

summary(mod3)

AIC(mod3)
```

Transformation 2 from project 1

```{r}
trans2 <- function(lambda,y){
    y.l <- 2*log((y^lambda)/(1-y)^(1-lambda))
    return(y.l)}

## profile likelihood for lambda
lp.lambda2 <- function(lambda,y){
    n <- length(y)
    y.l <- trans2(lambda ,y)
    sigmasq <- 1/n * sum((y.l-mean(y.l))^2)
    -n/2 * log(sigmasq) + sum(log(abs(((4*lambda - 2)*y - 2*lambda)/(y*(-1 + y)))))}
    
(opt.lambda.trans2=optimize(lp.lambda2,c(0.000001,1),y=data$pow.obs.norm, maximum=TRUE))
```
```{r}
mod4=glm(trans2(opt.lambda.trans2$maximum,data$pow.obs.norm)~ ws30,family=gaussian,  data=data)

summary(mod4)

AIC(mod4)

```

```{r}

## box-cox transformation
bc.trans <- function(lambda,y){
    y.l <- (y^lambda-1)/lambda
    
    if(lambda==0){y.l <- log(y)}
    return(y.l)}

## profile likelihood for lambda
lp.lambda <- function(lambda,y){
    n <- length(y)
    y.l <- bc.trans(lambda ,y)
    sigmasq <- 1/n * sum((y.l-mean(y.l))^2)
    -n/2 * log(sigmasq) + (lambda-1)*sum(log(y))}

(opt.lambda.boxcox=optimize(lp.lambda,c(-2,2),y=data$pow.obs.norm,maximum=TRUE))

```
```{r}
mod5=glm(bc.trans(opt.lambda.boxcox$maximum,data$pow.obs.norm)~ ws30,family=gaussian,  data=data)

summary(mod5)

AIC(mod5)
```

```{r}
library(betareg)
mod6=betareg(pow.obs.norm ~ ws30, data=data)

summary(mod6)

AIC(mod6)

qqnorm(mod6$residuals)
qqline(mod6$residuals)

#plot(mod6, which = 5)

#qqPlot(mod6, data = data)
```

```{r}


#qqnorm(dnorm(data$pow.obs.norm, mean = mod6$fitted.values, sd=mod6$)
       
```









```

